---
title: "Snow depth and SWE"
output: ioslides_presentation
date: '2022-05-31'
---

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

- In this short presentation we will look a few relationships between time, snow depth, and snow water equivalent (SWE) at various SNOTEL stations in Northwest Montana. 

- The SNOTEL stations we'll look at are Badger Pass, Emery Creek, and Pike Creek. 

- The data we use in this presentation was scraped from [this website](https://looper.avalanche.state.co.us/fac/), which aggregates snow related weather data for Northwest Montana. 

- The data we use was recorded at 6AM every day from October 1, 2021 until May 31, 2022. 

## Snow depth

```{r, loadData}
df <- read.csv("~/Desktop/Course 9 - Developing data products/DDP_projects/p2/flathead.csv") %>%
  filter( str_detect(Station, "Badger|Emery|Pike")) %>%
  select(c(2:4, 18, 19)) %>%
  mutate(date = as.Date(date, "%Y-%m-%d")) %>%
  na.omit()
```

First we look at snow depth over time, colored by station.

```{r, fig.align='center'}
p0 <- plot_ly(df, x = ~date, y = ~SnoHt, color = ~Station, 
              type = "scatter", mode = "line",
              text = ~paste("Elev: ", Elev, " ft.",
                            "<br>Date: ", date,
                            "<br>Depth: ", SnoHt, " in."),
              hoverinfo = "text") %>%
  layout( title = "Snow Depth: Oct '21 thru May '22",
          xaxis = list(title = ""),
          yaxis = list(title = "Snow Depth (in)"))
p0
```

## SWE over time, by station

Next we look at SWE over time, colored by station.

```{r, fig.align='center'}
p1 <- plot_ly(df, x = ~date, y = ~SWE, color = ~Station, 
              type = "scatter", mode = "line",
              text = ~paste("Elev: ", Elev, " ft.",
                            "<br>Date: ", date,
                            "<br>SWE: ", SWE, " in."),
              hoverinfo = "text") %>%
  layout( title = "SWE: Oct '21 thru May '22",
          xaxis = list(title = ""),
          yaxis = list(title = "SWE (in)"))
p1
```

## SWE vs snow depth

Finally we look at SWE vs snow depth, by station

```{r, fig.align='center'}
p2 <- plot_ly(df, x = ~SnoHt, y = ~SWE, color = ~Station, 
              type = "scatter", mode = "markers",
              text = ~paste("Elev: ", Elev, " ft.",
                            "<br>Date: ", date,
                            "<br>Depth: ", SnoHt, 
                            "<br>SWE: ", SWE 
                            ), 
              hoverinfo = "text") %>%
  layout(title = "SWE vs Depth: Oct 2021 thru May 2022",
         xaxis = list(title = 'Snow Depth (in)'),
         yaxis = list(title = 'SWE (in)'))
p2

```

## Some minor conclusions

- Badger pass, the station with the highest elevation, has the highest snow depth and SWE for the entire season. 

- Interestingly, Emery Creek is lower in elevation, but consistently has higher SWE and snow depth than Pike Creek. This may be because of aspect or station sun exposure.

- After each station hits its max SWE, the snow becomes more settled and dense, ie the same SWE corresponds to a more shallow snow pack. This explains the "loops" in each station's SWE vs Depth plot.  